The impacts of park access on health outcomes: A spatial comprehensive approach

### 2019-02-22 manuscript.Rnw
Revise literature review and data discussions

\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2}, Kari Watkins\affilnum{1}, 
and Professor WhogaveTweets\affilnum{3}}
\runninghead{Boyd, et al.}

\affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University\\
\affilnum{3}Somewhere at Georgia Tech}

\corrauth{Gregory Macfarlane\\
430 Engineering Building\\
Brigham Young University\\
Provo, UT 84602\\
USA}
\begin{abstract}
Recent research has underscored the potential for public green spaces to 
influence individual and societal health outcomes, but 
empirical measurements of such influences have yielded mixed results to date, with
particular disagreement surrounding how access to parks ought to be defined while
controlling alternate explanations. In this paper we present a comprehensive 
measure of park accessibility drawn from random utility choice theory, which avoids
arbitrary assertions of proximity while incorporating potentially numerous 
amenities and attributes of both the parks and the population. We apply this 
choice-based accessibility measure to Census tract-level health and 
socioeconomic data in New York City, paired with geographic parks data and  
and geolocated data collected from the Twitter API. Controlling for the 
socioeconomic variables and spatially correlated error terms, we show a positive
but suggestive relationship between park access and physical activity rates, and
a clear negative relationship between park access and obesity rates.
\end{abstract}

\maketitle
<<setup, cache=FALSE, echo=FALSE, message=FALSE>>=
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(RSocrata)
library(tidycensus)

@@

The United States is currently facing an epidemic of obesity and chronic
diseases, which are non-communicable diseases of long duration and typically
slow development, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer.
Recent statistics suggest that approximately 300,000 premature deaths each year
can be attributed to chronic diseases \citep{WHO2014}.
According to \citet{Finkelstein2009}, obese Americans spend approximately
\$1,500 more in healthcare costs annually than Americans of
average weight, totaling \$147 billion in direct medical expenses nationally.

@@

Until recently, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise. 
According to \citet{Coutts2008} however, this trend is shifting as professionals
“have begun adopting an ecological paradigm, accepting that both individual and
environmental determinants play a role in health behavior. This new, arguably
revisited, public-health paradigm accounts not only for the compositional (who
you are) but also for the contextual (where you are) influences on physical
activity.” As professionals begin to operate from the assumption that the design
and configuration of the built environment can facilitate or inhibit physical
activity, they are increasingly looking to public spaces like parks and
greenways as key elements of the built environment that can support physical 
activity \citep{Bedimo-Rung2005a}.

In spite of this interest and the numerous studies attempting to empirically 
estimate the strength of such support, evidence to date has been mixed. 
A major challenge researchers have faced is a proliferation of techniques for
measuring access to green spaces, of which many are plausible but lack 
analytical or theoretical basis. In this paper, we present a holistic 
and flexible measurement for park accessibility based in random utility 
choice theory. This choice-based accessibility measure compares the continuous
distance to all parks in the region, weighted against the size of the park and 
its other amenities. We apply this measurement to study the link between park
accessibility and attractiveness and Census tract-level aggregate physical
activity participation and obesity rates in New York City, controlling for
spatial correlation and tract-level socioeconomic characteristics. 
We find a positive relationship where the
least park-accessible tracts have an expected physical activity participation
rate half a percentage point lower than the most accessible tracts, though 
this result is only suggestive of correlation from a statistical perspective.
However, we also find a strong and significant correlation between 
increased park accessibility and decreased obesity rates, again controlling for
spatial correlation, socioeconomic attributes, and physical activity rates.

We also demonstrate how the choice-based accessibility metric could potentially 
be expanded to account for other elements of a park's attractiveness by 
introducing the number of tweets emanating from within a park during 
(WE NEED TO DESCRIBE THE PERIOD). The paper concludes with a discussion of
opportunities for future research.

@@

\begin{table*}
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{11cm}}
\toprule
Category & Summary of Benefits\\
\midrule
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce heat-related mortality \citep{Stone2006}\\
 & Reduce incidence of cardiovascular-related mortality \citep{Mitchell2008}\\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
\midrule
Mental Health & Reduce stress and mental fatigue \citep{Wolch2014}\\
  & Reduce aggression \citep{Kuo2001a}\\
  & Enhance emotional and cognitive development\\
  & Improve behavioral outcomes in youth \\
\midrule
Social Health & Enhanced community aesthetics\\
 & Reduce crime \citep{Kuo2013,Wolfe2012}\\
 & Increase social interaction \citep{Sullivan2004}\\
\midrule
Economic Health & Provide ecosystem services\\
 & Increase residential property values and municipal property tax revenues\\
 & Attract more shoppers and increase economic activity to commercial districts \\
\bottomrule
\end{tabular}
\end{table*}

In spite of these generally positive findings, the evidence of a specific link
between green space and physical activity is somewhat mixed.
\citet{Wolf2008} reports that people self-identify as more active and attain a 
higher quality of life when a greater portion of their environment is greenspace.
\citet{Larson2016} similarly used self-reported scores on the Gallup-Healthways Wellbeing
Index to evaluate the relationship between different areas of wellbeing,
including physical health, and park quantity, quality, and accessibility in 44
U.S. cities. While the authors found positive associations between wellbeing
and park quality and accessibility, these relationships were not statistically
significant. A study by \citet{West2012} used park data from the
Trust for Public Land’s 2010 City Park Facts and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
between the density of parkland, parkland per capita, and levels of physical
activity and obesity for 67 metropolitan statistical areas in the U.S. The study
found a significant, positive association between park density and physical
activity and a significant, negative association between park density and
obesity. Conversely, \citet{Richardson2012} examined the
relationship between urban greenspace and selected mortality rates, and though the authors
did not find a statistically significant relationship between the quantity of
urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
car accidents, they did find that all-cause mortality was significantly higher in
greener cities. In a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured physical
activity, \citet{Bancroft2015} found that five studies reported a significant
positive association between the two, six studies produced mixed results, and
nine studies found no association at all.

While several of the aforementioned studies reported results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, the large majority
of extant research has been conducted at the level of the city or 
metropolitan statistical area.
Metropolitan-level analyses do not capture the within-city variation in
accessibility to parks that exist in many cities. Some cities with large amounts
of total greenspace may nevertheless inequally distribute this space throughout
the city, leading to areas with poor access. Conversely, a city with a smaller
overall proportion of greenspace may give all of its citizens better access to
sufficient greenspace to meet their needs or wants for physical activity.
Considering access at the neighborhood level within a single city may also
eliminate some regional or cultural fixed effects affecting metropolitan-level
analyses.

In a study of neighborhoods in New York City, \citet{Stark2014} found that a
higher proportion of park space in a neighborhood was associated with a lower
BMI for its residents. \citet{Pretty2005} showed the presence of trees and other
vegetation in outdoor environments is positively associated with physical
activity; further, \citet{Suminski2005} showed that adding vegetation to
existing sidewalks and trails increased local motivation to engage in physical
activity.

@@

Importantly, \citeauthor{Vale2016} acknowledge that there is not yet a
consensus on the most appropriate accessibility metric to use in a given
setting, noting that ``ways to measure active accessibility are as varied as the
number of scholars the measure them.''(p. 227) This is at least partly because
each method has inherent strengths and weaknesses.
Distance is an inherently continuous phenomenon, but software tools ease and enable
buffer analyses where distance becomes a binary phenomon of somewhat arbitrary
extent. \citet{Logan2017} note that these commonly used approaches can also lead
researchers to overlook access-poor populations.
Infrastructure-based distances can vary widely
depending on which modes of travel are considered: a highway network may not
adequately capture trails and bike paths.
Similarly, gravity-based measures may not always account for surrounding land
uses when considering the attractiveness or impedance of a destination; the
findings of \citet{Pretty2005} and others suggest unpleasant pedestrian routes
attenuate perceived travel costs.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus to individual $i$, or the total weighted value of all 
alternatives in individual $i$'s choice set, can
be obtained as the logarithm of the sum in the denominator of the choice 
probability equation \citep{Small1981}:
\begin{equation}\label{eq:cs}
C_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to such a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher. 
This allows for the fact that some people are more or less sensitive to 
distances, and that distance is a continuous, non-binary phenomenon. 
It defies reason to assume person living 1.1 miles 
from a park has functionally different accessibility than someone living at 0.9 miles.
Second, the random utility formulation
allows the researcher to include any attribute of the park as part of its utility
specification; in this case, we consider the size of the park weighted against its
distance as an element of relative accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park on Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

Note that this formulation is an extension of the gravity-based accessibility
statistics, in the same way that the gravity-based trip distribution model model 
is a specific case of
a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
allows us to consider more attributes of the park or the indivdual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size, and the
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

In spite of its flexibility and basis in choice theory, accessibility measures 
based in log-sums
have not received much application in the accessibility literature compared with
distance-based or even
gravity-based measures. \citet{Zhang2011} present a metric with a similar functional
form, though the authors arrive at it from a motivation other than destination 
choice theory. Log-sums are commonly used, however, in alternatives analyses of
transit infrastructure improvements \citep{DeJong2007}. A reason for this is
likely that a regional travel demand model is available to the analysts, thus making 
calibrated and multimodal logsums are readily available \citep{Geurs2010}.

\section{Empirical Application}
In this section we describe an experiment with data for New York City, where
we compute a holistic logsum-based accessibility to parks metric and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors. We subsequently model the effect
of this accessibility on obesity.

@@

@@

@@

This study uses data available to the public from a variety of federal and state
data agencies. The datasets, as well as the analysis code, are available on GitHub
at \url{https://github.com/boydnpb/park_access_new.git}.
The primary dataset is a geographic polygons shapefile of Census tracts in
New York City. We appended relevant sociodemographic data for each tract level
from the American Community Survey 2011-2015 5-year estimates. Data on physical
activity participation rate and obesity rates for each tract are available from
the Centers for Disease Control and Prevention’s 500 Cities Project (WE NEED A CITATION).
After removing tracts with missing population information or outlying and unusual
characteristics, we have \Sexpr{nrow(tracts)} complete cases. Table
\ref{tab:tractsdata} presents key descriptive statistics for these data.

@@

In a destination choice framework, the tracts represent the ``origins'' and
``destinations'' are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance (WE NEED A CITATION).
Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space. This
leaves us with \Sexpr{nrow(parks)} parks. It is worth noting that the particular
dataset classifies specific facilities within larger parks as independent
features: baseball fields within Central Park, for example, are themselves
delineated green spaces. We elect to retain these independent features as they may
generate additional physical activity participation aside from the containing 
park.

>>>>

\author{Nico Boyd\affilnum{1}, Gregory S. Macfarlane\affilnum{2},
John Taylor\affilnum{1}, and Kari Watkins\affilnum{1}}
\runninghead{Boyd, et al.}

\affiliation{\affilnum{1}School of Civil and Environmental Engineering, Georgia Institute of Technology\\
\affilnum{2}Civil and Environmental Engineering Department, Brigham Young University}

\corrauth{Gregory Macfarlane\\
430 Engineering Building\\
Brigham Young University\\
Provo, UT 84602\\
USA}
\begin{abstract}
Recent research has underscored the potential for public green spaces to
influence individual and societal health outcomes, but
empirical measurements of such influences have yielded mixed results to date, with
particular disagreement surrounding how access to parks ought to be defined while
controlling alternate explanations. In this paper we present a comprehensive
measure of park accessibility drawn from random utility choice theory, which avoids
arbitrary assertions of proximity while incorporating potentially numerous
amenities and attributes of both the parks and the population. We apply this
choice-based accessibility measure to Census tract-level health and
socioeconomic data in New York City, paired with geographic parks data and
geolocated position data collected from the Twitter API. Controlling for the
socioeconomic variables and spatially correlated error terms, we show a positive
but suggestive relationship between park access and physical activity rates, and
a clear negative relationship between park access and obesity rates.
\end{abstract}

\maketitle
<<setup, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE>>=
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(RSocrata)
library(tidycensus)

@@

The United States is currently facing an epidemic of obesity and chronic
diseases, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer.
Recent statistics suggest that approximately 300,000 premature deaths each year
can be attributed to chronic diseases \citep{WHO2014}.
According to \citet{Finkelstein2009}, obese Americans spend approximately
\$1,500 more in healthcare costs annually than Americans of
average weight, totaling \$147 billion in direct medical expenses nationally.

@@

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
According to \citet{Duhl1999} however, the field of public health has
moved toward a new paradigm that is grounded in a social model of health,
treating health ``as an outcome of the effects of socioeconomic status, culture,
environmental conditions, housing, employment and community influences.''
Further, \citet{Coutts2008} maintains that awareness of the impacts of the built
environment on health will increase connections between the practices of public
health and urban planning. As professionals from both fields begin to operate
from the assumption that the design and configuration of the built environment
can facilitate or inhibit physical activity, they are increasingly looking to
public spaces like parks and greenways as key elements of the built environment
that can support physical activity \citep{Bedimo-Rung2005a}.

In spite of this interest and the numerous studies attempting to empirically
estimate the strength of such support, evidence to date has been mixed.
A major challenge researchers have faced is a proliferation of techniques for
measuring access to green spaces, of which many are plausible but lack
analytical or theoretical basis. In this paper, we present a holistic
and flexible measurement for park accessibility based in random utility
choice theory. This choice-based accessibility measure compares the continuous
distance to all parks in the region, weighted against the size of the park and
its other amenities. We apply this measurement to study the link between park
accessibility and attractiveness and Census tract-level aggregate physical
activity participation and obesity rates in New York City, controlling for
spatial correlation and tract-level socioeconomic characteristics.
We find a positive relationship where the least park-accessible tracts have an
expected physical activity participation rate half a percentage point lower than
the most accessible tracts, though this result is only suggestive of correlation
($p$-value $<0.1$ from a statistical) perspective.
However, we also find a strong and significant correlation between
increased park accessibility and decreased obesity rates, again controlling for
spatial correlation, socioeconomic attributes, and physical activity rates.

We also demonstrate how the choice-based accessibility metric could potentially
be expanded to account for other elements of a park's attractiveness by
introducing the number of geolocated tweets emanating from within each park during
September 2014. The paper concludes with a discussion of
opportunities for future research.

@@

\begin{table*}
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{11cm}}
\toprule
Category & Summary of Benefits\\
\midrule
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce heat-related mortality \citep{Stone2006}\\
 & Reduce incidence of cardiovascular-related mortality \citep{Mitchell2008}\\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
\midrule
Mental Health & Reduce stress and mental fatigue \citep{Wolch2014}\\
  & Reduce aggression \citep{Kuo2001a}\\
  & Improve behavioral outcomes in youth \citep{Taylor1998, Madzia2018} \\
\midrule
Social Health & Reduce crime \citep{Wolfe2012}\\
 & Increase social interaction \citep{Sullivan2004}\\
\midrule
Economic Health & Increase residential property values and municipal property tax revenues \citep{Donovan2010}\\
 & Attract more shoppers and increase economic activity to commercial districts \citep{Schwab2009}\\
\bottomrule
\end{tabular}
\end{table*}

In spite of these generally positive findings, the evidence of a specific link
between green space and physical activity is somewhat mixed.
\citet{Wolf2008} reports that people self-identify as more active and attain a
higher quality of life when a greater portion of their environment is greenspace.
\citet{Larson2016} similarly used self-reported scores on the Gallup-Healthways Wellbeing
Index to evaluate the relationship between different areas of wellbeing,
including physical health, and park quantity, quality, and accessibility in 44
U.S. cities. While the authors found positive associations between wellbeing
and park quality and accessibility, these relationships were not statistically
significant. A study by \citet{West2012} used park data from the
Trust for Public Land’s 2010 City Park Facts and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
between the density of parkland, parkland per capita, and levels of physical
activity and obesity for 67 metropolitan statistical areas in the U.S. The study
found a significant, positive association between park density and physical
activity and a significant, negative association between park density and
obesity. Conversely, \citet{Richardson2012} examined the
relationship between urban greenspace and selected mortality rates, and though the authors
did not find a statistically significant relationship between the quantity of
urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
car accidents, they did find that all-cause mortality was oddly, {\em higher} in
greener cities. In a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured physical
activity, \citet{Bancroft2015} found that five studies reported a significant
positive association between the two, six studies produced mixed results, and
nine studies found no association at all.

While several of the aforementioned studies reported results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, the large majority
of extant research has been conducted at the level of the city or
metropolitan statistical area.
Metropolitan-level analyses do not capture the within-city variation in
accessibility to parks that exist in many cities. Some cities with large amounts
of total greenspace may nevertheless unequally distribute this space throughout
the city, leading to areas with poor access. Conversely, a city with a smaller
overall proportion of greenspace may give all of its citizens better access to
sufficient greenspace to meet their needs or wants for physical activity.
Considering access at the neighborhood level within a single city may also
eliminate some regional or cultural fixed effects affecting metropolitan-level
analyses.

In a study of neighborhoods in New York City, \citet{Stark2014} used
hierarchical linear models to explore relationships between the proportion of
zip codes that was park space, the cleanliness of parks, and the body mass index
of adult residents. The authors found that a
higher proportion of park space in a neighborhood was associated with a lower
BMI for its residents, and lower park cleanliness scores were associated with a
higher BMI, suggesting that increasing the supply of clean, well-maintained
parks could positively impact adult BMI.
\citet{Suminski2005} showed that adding vegetation to
existing sidewalks and trails increased local motivation to engage in physical
activity.

@@

Importantly, \citeauthor{Vale2016} acknowledge that there is not yet a
consensus on the most appropriate accessibility metric to use in a given
setting, noting that ``ways to measure active accessibility are as varied as the
number of scholars that measure them.''(p. 227) This is at least partly because
each method has inherent strengths and weaknesses.
Distance is an inherently continuous phenomenon, but software tools ease and enable
buffer analyses where distance becomes a binary phenomenon of somewhat arbitrary
extent, as a cut-off distance is implied in the buffer distance assumption.
\citet{Logan2017} note that these commonly used approaches can also lead
researchers to overlook access-poor populations.
Infrastructure-based distances can vary widely
depending on which modes of travel are considered: a highway network may not
adequately capture trails and bike paths.
Similarly, gravity-based measures may not always account for surrounding land
uses when considering the attractiveness or impedance of a destination; the
findings of \citet{Pretty2005} and others suggest unpleasant pedestrian routes
attenuate perceived travel costs.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus to individual $i$, or the total weighted value of all
alternatives in individual $i$'s choice set, can
be obtained as the logarithm of the sum in the denominator of the choice
probability equation \citep{Small1981}:
\begin{equation}\label{eq:cs}
C_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to such a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher.
This allows for the fact that some people are more or less sensitive to
distances, and that distance is a continuous, non-binary phenomenon.
It defies reason to assume a person living 1.1 miles
from a park has functionally different accessibility than someone living at 0.9 miles.
Second, the random utility formulation
allows the researcher to include any attribute of the park as part of its utility
specification; in this case, we consider the size of the park weighted against its
distance as an element of relative accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park in Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

Note that this formulation is an extension of the gravity-based accessibility
statistics, in the same way that the gravity-based trip distribution model
is a specific case of
a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
allows us to consider more attributes of the park or the individual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size, and the
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

In spite of its flexibility and basis in choice theory, accessibility measures
based in log-sums
have not received much application in the accessibility literature compared with
distance-based or even
gravity-based measures. \citet{Zhang2011} present a metric with a similar functional
form, though the authors arrive at it from a motivation other than destination
choice theory. Log-sums are commonly used, however, in alternatives analyses of
transit infrastructure improvements \citep{DeJong2007}. A reason for this is
likely that a regional travel demand model is available to the analysts, thus making
calibrated and multimodal logsums readily available \citep{Geurs2010}.

\section{Empirical Application}
In this section, we describe an experiment with data for New York City, where
we compute a holistic logsum-based accessibility to parks measure and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors. We subsequently model the effect
of the accessibility metric on obesity rates, accounting for physical activity and
the other controls. We then model the relationship between a modified logsum
considering geolocated Twitter data and physical activity and obesity rates.

@@

This study uses data available to the public from a variety of federal and state
data agencies. The datasets, as well as the analysis code, are available on GitHub
at \url{https://github.com/boydnpb/park_access_new}.

@@

The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program. The models
are estimated on Census sociodemographic data paired with national and local
health surveys, and have been validated at multiple geographic levels
\citep{Wang2018, Wang2017, Zhang2015}. We use two synthetic indicators as our
outcome variables in this study: the share of adults in a Census tract who are
obese, and the share of adults who participate in no leisure-time physical activity.
To improve clarity in our interpretation, we use the complement of the second variable ---
the share of tract adults who participate in {\em some} physical activity ---
even if the amount is not sufficient to affect overall health. Both indicators
are estimated for the year 2016.

@@

To the health data, we join data collected through the Census Bureau API via
the {\texttt tidycensus} package for R \citep{Walker2019}. The primary dataset
is a geographic polygons layer of Census tracts in the five boroughs of New York
City. We append to each Census tract relevant sociodemographic data
for each tract level from the American Community Survey 2013-2017 5-year
estimates. For a small handful of tracts in our sample, Census supressed the
median income estimate; these appear to be primarily wealthy tracts and in
almost all cases the CDC estimates of obesity and physical activity are missing
as well. After removing these tracts from the estimation dataset, we have
\Sexpr{nrow(tracts)} complete cases. Table \ref{tab:tractsdata} presents key
descriptive statistics for these data.

@@

In a destination choice framework, the tracts represent the ``origins'' and
``destinations'' are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance \citep{nycparks}.
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium and its surrounding
parking lots. We also removed parks of less than half an acre in size, as these
appear to be predominantly vacant lots or parking strips rather than legitimate
public green space. This leaves us with \Sexpr{nrow(parks)} parks. It is worth
noting that the particular dataset classifies specific facilities within larger
parks as independent features: baseball fields within Central Park, for example,
are themselves delineated green spaces. We elect to retain these independent
features as they may generate additional physical activity participation aside
from the containing park. Additionally, large cemeteries are not considered
open space in these data, though people may use the space for physical activities
such as walking or running. For each park we calculate its size in acres, and
identify the number of geolocated tweets emanating from the park in September
2014, as a proxy for the park's use or popularity.
